\chapter*{Acknowledgments / Remerciements}
\addcontentsline{toc}{chapter}{Acknowledgments / Remerciements}

First I would like to thank the board of Doctoral School 352 that granted me this PhD funding, Andrea Marini for hosting me one week in Rome, funded by the Cost Action TUMIEE, Daniele Varsano and HPC Europa3 for allowing me a 2-month visit in Modena.

I would like to thank my supervisor Claudio for his constant support, his availability, for giving me the freedom to study the things I wanted and for his wise decisions that helped me to become a scientist. Un ringraziamento speciale per il tempo che hai dedicato alla revisione di questo manoscritto.

I thank Fulvio for the time he dedicated to train me on different topics and for his hospitality in Roma and Modena. 

I would like to thank Khaoula Boukari and Andres Saul for the management and maintenance of the computer cluster \textit{Rosa} which was an undeniable advantage for this PhD work.

I thank the TSN department as a whole for scientific discussions and at large, everyone who I discussed about science with and who sated my hunger for understanding, Davide Sangalli, Elena Cannuccia, Adrien Rousseau, Thibaut Sohier, Pedro Melo, Martino Silvetti, Matteo Zanfrognini, Riccardo Reho.

I thank the people I met in CINaM and who became my friends : Anna, Petr, Clémence, Céline, Alex, Julien, Diego, Ilan, Rajarshi, Guillem, Djouher, Mathilde, Kostis, Tarik, Gaëlle, Giorgia, Luigi,  Mohit, Vyshnav. You contributed to brighten the sometimes gloomy days. 

I thank Vitaly, Laura, Miguel and Kalyani for their help in the organization of YRM 2022 which was quite an experience, Véronique, Régine, Lydia, Cindy, Sabina for the precious administrative help, and Christophe for the budgetary rescue.

I thank Marine Baujon for helping me manage things that felt bigger than me at times.

I thank everyone for good times in Marseille and elsewhere Sam, Yann, Jonathan, Pavlo, Vasile, Thaïs, Simon, Inès.

Merci à mes proches qui donnent un sens à tout ceci. Jules, Pieric, Valentin, Yann, merci de votre soutien indéfectible, vous êtes des amis formidables.

Un simple merci ne suffirait pas pour exprimer ma gratitude envers mes parents et ma famille. Merci Maman et Papa pour m'avoir toujours soutenu et exprimé votre fierté. Merci Baptiste et Maxime d'être des frères admirables. Merci Côme et Thaïs de bien vouloir jouer avec moi.